---
title: 'Coursera_Peer-graded Assignment: Prediction Assignment Writeup'
author: "Sanoop"
date: "13/05/2020"
output: html_document
---

#Overview
The goal of this project is to predict the manner in which they did the exercise.

#Running the Libraries
```{r}

library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
```

#Reading the datasets
```{r}
training <- read.csv("C:/Users/sanoo/OneDrive/Desktop/Coursera/pml-training.csv")
testing <- read.csv("C:/Users/sanoo/OneDrive/Desktop/Coursera/pml-testing.csv")
```

#Data processing
```{r}
library(caret)
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
dim(TestSet)
```

# remove the variables with Nearly zero Variance
```{r}
n0var <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -n0var]
TestSet  <- TestSet[, -n0var]
dim(TrainSet)
dim(TestSet)
```

# remove variables that are mostly NA
```{r}
AllNA <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet <- TestSet[, AllNA==FALSE]
dim(TrainSet)
dim(TestSet)
```

# remove identification only variables (columns 1 to 5)
```{r}
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
dim(TestSet)
```

#checking correlation among variables
```{r}
library(corrplot)
M <- cor(TrainSet[, -54])
corrplot(M, method="circle")
```

#Random forest method
# model fit
```{r}
set.seed(3408)
controlRF <- trainControl(method="cv", number=3, verboseIter=FALSE)
modFitRandForest <- train(classe ~ ., data=TrainSet, method="rf", trControl=controlRF)
modFitRandForest$finalModel
```

# prediction on Test dataset
```{r}
predictRandForest <- predict(modFitRandForest, newdata=TestSet)
confMatRandForest <- confusionMatrix(predictRandForest, TestSet$classe)
confMatRandForest
```

# plot matrix results
```{r}
plot(confMatRandForest$table, col = confMatRandForest$byClass, 
main = paste("Random Forest - Accuracy =", round(confMatRandForest$overall['Accuracy'], 4)))
```

#Applying the selected Model to the Test Data (Random Forest is selected because of higher accuracy)
```{r}
predictTEST <- predict(modFitRandForest, newdata=testing)
predictTEST
```










